---
title: 'HW1'
output: 
  html_document:
    theme: flatly
link-citations: yes
---

# Step 1: Loading and preliminary checks


## Loading the data

Installing the necessary packages

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
#install.packages(c('data.table','leaflet'))
library(data.table)
library(leaflet)
library(tidyverse)
```

```{r check-file, include=T}
fn <- 'https://raw.githubusercontent.com/dungwoong/JSC370HW1/master/Bicycle_Thefts.csv'
if (!file.exists('Bicycle_Thefts.csv'))
  download.file(fn, destfile = 'Bicycle_Thefts.csv')
df <- data.table::fread('Bicycle_Thefts.csv')
```

## Preliminary checks
The dataset description can be found [here](https://ago-item-storage.s3.us-east-1.amazonaws.com/c0b17f1888544078bf650f3b8b04d35d/PSDP_Open_Data_Documentation.pdf?X-Amz-Security-Token=IQoJb3JpZ2luX2VjEKn%2F%2F%2F%2F%2F%2F%2F%2F%2F%2FwEaCXVzLWVhc3QtMSJIMEYCIQDj5ufbbDJ%2Bn3Lbrsac6%2FJSfYwygHN6QJgRbz5R31R%2BJgIhAIaxKulHgpveGrvOg1U1uOyuUweOhsVDN%2BYVMNq1VgJaKswECCIQABoMNjA0NzU4MTAyNjY1IgxH%2BxY4ffK1IOFrQfgqqQSUq8iEVvqvHK327wJ6J6AFQiWPDzWBKx%2FbkDpZcu8Zkh5nOsBraV7DHrW2He3JDPrq%2Fop3vAlyDkCKNjqFAx05IRXE6YpXCrylH2yoNfWBw2PJ2O0k5NAcaLN66a31f2Qc6obOy%2FQdRzU0n4KbAPYZH7GEuwlsKUNtYMQdoBlnbUcxRkIMRbF%2BUiu5Z5Jn%2B6Ym5d9DMG%2FGM1P13zA%2BsdJJsurPGz%2BjgnwAovoygfj9nUcWzLnjpAEnrf0rMaHxg04D%2B0U%2FkRS9soRZHqYyG3wAHd9M7kAJIehYHvntX9ihwdOgu6MYorKznUdX4rUgzHmY9ujaiFP9HEhHLAxcy7ZDMdAEFxgidFFnobo5thV%2BSCPmhV9hEJ2xdRFIosCI8Qh%2B1iA8M5bkOc8z2OTIFButSQgA4rjwtsB5TGKboxKZz5aiIo0EWdUCkBuYBy0nB8qBmkaKKwOuIelrAI4%2BoMpYWh5cR5YohNcPVLq3yV4ORDowHSuo96e33vUdrQIGl6pkyQpedUp6s%2BpJ0UA2evMxuGPXcZTZPh2IhpxyQ6nPECFIRbT5lgRl37vpGKCi7hdhXlwG6%2FDz7q8bPHJWWjZyXgkxF%2FUwFpgUIwX6oXIpezKm0auoqdsgmv0ie0bg9TbZU4TuzUY0bn1J6uuKhAhlqt58i3Nprpf85zrY%2BcQCWzwEm7Hs%2FbDBygfKPeyU4Tbc8ro8W1AXCWjcJDndY%2BXrPUtYu4BWJ4tWML6ax54GOqgBtKNVCJ6W33wc7Fn9mcmTWw7Y6sfIOZPAkKMC%2FC0J1L2HXzr4ww%2B%2FRU3g7mNB%2B37Sb9llvl9hbefcKvnhHHmvN7AdakxDRPBG7C8nP%2B2u%2FHJc52gX%2FIVgbxMS%2Bim4Iq6DBEm9zpm2gtHRmHEM2H%2FFuqcLh9PeGbE6Mg28O64yjDkIy6%2Bfv7GwAGay%2F312gT%2BjFRkcLfkggzCLPnlXmQDFtU3DiZojLFrE&X-Amz-Algorithm=AWS4-HMAC-SHA256&X-Amz-Date=20230126T020707Z&X-Amz-SignedHeaders=host&X-Amz-Expires=300&X-Amz-Credential=ASIAYZTTEKKE6QZBTDYN%2F20230126%2Fus-east-1%2Fs3%2Faws4_request&X-Amz-Signature=8da3bd61eca98818b1d57f7dc9a2249abdf01dc883471fb1be2d5a2012a1b370)

The dataset should have 30154 records. The table should have 32 attributes, although the CSV downloads with 35 columns.
```{r}
dim(df)
```

We're only interested in a few variables for this investigation. Specifically, all time variables concerning the reporting time and location of the theft. 

These variables are listed under different names than that specified in the data description document, but we can infer their meanings. I will decide on variables to investigate, then check for missing or nonsensical values in the variables I want to investigate.

## Variables to investigate

The Primary_Offence column lists many crimes, but the data portal states that "(all) occurrences are related to...offences where the theft of a bicycle was included," so I will assume it's not necessary and I can assume each row is related to a bicycle theft.

Since we are investigating the location and time of reports, I will discard variables regarding the time/location of the occurrence of each incident.

I will also discard other bike metadata and statistics relating to the crime, such as bike cost, status of bike(stolen/recovered/unknown), etc.

I will also discard the 'X' and 'Y' variables in the CSV, as they are not explained in the data description and do not make sense to me. I will also discard any id variables.
```{r, result='hide', fig.show='hide', message=F}
# this code was used to check the other variables
# str(df)
# summary(df)

# table(df$Primary_Offence)

# table(df$Status)
```

Overall, there are a few variables I'm interested in.
These are listed below:

Name in Dataset | Description
-----|------------
Report_* (multiple variables) | Time variables regarding reporting
Hood_ID | Identifier of each neighborhood
NeighbourhoodName | Name of neighborhood
City | Name of City
Longitude | Longitude coordinates(offset to nearest intersection)
Latitude | Latitude coordinates(offset to nearest intersection)

I checked each of these variables and verified that no loading errors occurred. In the next section, I will remove the other variables and perform further checks

# Step 2: Clean the data

Report_Date is reflected in the other report variables, and the time of the report dates are all 4:00:00, suggesting a loading error. Report_Hour most likely reports the correct hour of the report. Thus, we will remove Report_Date
```{r}
feats <- c('Report_Date', 'Report_Year', 'Report_Year', 'Report_Month', 'Report_DayOfMonth', 'Report_Hour')
df %>% select(feats) %>% head(5)
```


```{r}
feats <- c('Report_Year', 'Report_Month', 'Report_DayOfWeek', 'Report_DayOfMonth', 'Report_DayOfYear', 'Report_Hour', 'City', 'Hood_ID', 'NeighbourhoodName', 'Longitude', 'Latitude')
df <- df %>% select(feats)
summary(df)
```
There seems to be some 0 values for Longitude and latitude. The summary shows that all hours, days of year, days of month and year are valid.

Among these variables, there are no null values.
```{r}
colSums(is.na(df))
```
TODO turn categoricals into categoricals, look for more outliers.